### 2014-10-22 fundamentals/examples.tex
tables where I put them
reduces the chaos in the bibliographies to just one file of chaos

A number of projects and tools outside of the core simulation kernel have been 
developed which improve the scope and the diversity of the capabilities in the \Cyclus 
ecosystem. Table \ref{tab:coretools} lists the tools and projects developed 
under close integration with the \Cyclus kernel.
\begin{table}
\centering
\begin{tabularx}{\textwidth}{|r|X|r|}
\hline
\textbf{Name} & \textbf{Description} & \textbf{Citation} \\
\hline
Cycic &  Input control - embedded in Cyclist & \cite{cycic}\\
Cyclist & Interactive data exploration environment & \cite{cyclist} \\
Ciclus & Continuous Integration scripts for \Cyclus & \cite{ciclus}\\
Cycstub & Skeleton for clean slate module development & \cite{cycstub}\\
Cyan & \Cyclus analysis tool & \cite{cyan}\\
Cloudlus & Tools for running \Cyclus in a cloud environment & \cite{cloudlus} \\
\hline
\end{tabularx}
\caption{Many tools outside of the scope of the \Cyclus kernel have been 
developed for improved user, developer, and analyst experinces with \Cyclus.}
\label{tab:coretools}
\end{table}

@@

Many archetypes external to the \Cycamore library  
have been \cite{huff_cyder_2013,,streamblender,commodconverter} 
and are being \cite{britelite,utk,mktdriveninst} developed for contribution to the 
\Cyclus ecosystem. These archetypes provide the first examples of 
developer-contributed capabilities.

\begin{table}
\centering
\begin{tabularx}{\textwidth}{|r|X|r|}
\hline
\textbf{Name} & \textbf{Description} & \textbf{Citation} \\
\hline
Brite-lite & A physics-based reactor archetype and fuel fabrication archetype & \cite{brightlite} \\
Nuclear Fuel Inventory Model & A flexible, ORIGEN-based, reactor analysis module & \cite{NFIM} \\
CommodConverter & A simple commodity converting storage facility archetype  & \cite{commodconverter} \\
MktDrivenInst & An institution that controls deployment based on commodity availability & \cite{mktdriveninst} \\
SeparationsMatrix & A facility for elemental separations of used fuel & \cite{streamblender} \\
StreamBlender & A facility for fuel fabrication from reprocessed streams & \cite{streamblender} \\
\hline
\end{tabularx}
\caption{A diverse set of archetypes under development reflect the diverse 
needs of researchers at various institutions. These archetypes, contributed 
outside of the \Cyclus core and \Cycamore libraries are the first demonstration 
of community-driven development in a fuel cycle simulator.}
\label{tab:archetypes}
\end{table}

>>>>

A number of projects and tools outside of the core simulation kernel have been 
developed which improve the scope and the diversity of the capabilities in the \Cyclus 
ecosystem. Table \ref{tab:coretools} lists the tools and projects developed 
under close integration with the \Cyclus kernel.
\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|r|X|r|}
\hline
\textbf{Name} & \textbf{Description} & \textbf{Citation} \\
\hline
Cycic &  Input control - embedded in Cyclist & \cite{flanagan_input_2013}\\
Cyclist & Interactive data exploration environment & \cite{livnat_cyclist_2014} \\
Ciclus & Continuous Integration scripts for \Cyclus & \cite{scopatz_ciclus_2014}\\
Cycstub & Skeleton for clean slate module development & \cite{carlsen_cycstub_2014}\\
Cyan & \Cyclus analysis tool & \cite{carlsen_cyan_2014}\\
Cloudlus & Tools for running \Cyclus in a cloud environment & \cite{carlsen_cloudlus_2014} \\
\hline
\end{tabularx}
\caption{Many tools outside of the scope of the \Cyclus kernel have been 
developed for improved user, developer, and analyst experinces with \Cyclus.}
\label{tab:coretools}
\end{table}

@@

Many archetypes external to the \Cycamore library  
have been 
\cite{huff_streamblender_2014,huff_commodconverter_2014}
and are being 
\cite{flanagan_bright-lite_2014,skutnik_nuclear_2014,huff_mktdriveninst_2014} developed for contribution to the 
\Cyclus ecosystem. These archetypes provide the first examples of 
developer-contributed capabilities.

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|r|X|r|}
\hline
\textbf{Name} & \textbf{Description} & \textbf{Citation} \\
\hline
Brite-lite & A physics-based reactor archetype and fuel fabrication archetype & \cite{flanagan_bright-lite_2014} \\
Nuclear Fuel Inventory Model & A flexible, ORIGEN-based, reactor analysis module & \cite{skutnik_nuclear_2014} \\
CommodConverter & A simple commodity converting storage facility archetype  & \cite{huff_commodconverter_2014} \\
MktDrivenInst & An institution that controls deployment based on commodity availability & \cite{huff_mktdriveninst_2014} \\
SeparationsMatrix & A facility for elemental separations of used fuel & \cite{huff_streamblender_2014} \\
StreamBlender & A facility for fuel fabrication from reprocessed streams & \cite{huff_streamblender_2014} \\
\hline
\end{tabularx}
\caption{A diverse set of archetypes under development reflect the diverse 
needs of researchers at various institutions. These archetypes, contributed 
outside of the \Cyclus core and \Cycamore libraries are the first demonstration 
of community-driven development in a fuel cycle simulator.}
\label{tab:archetypes}
\end{table}

### 2014-10-22 fundamentals/intro.tex
reduces the chaos in the bibliographies to just one file of chaos

An alternative to these approaches is a dynamic, agent-based simulator with an 
open platform and an ecosystem of research-driven capabilities.  Modern 
insight indicates that dynamic nuclear fuel cycle analysis more realistically 
supports a broader range of simulation goals than static analysis 
\cite{piet_dynamic_2011}. Similarly, experience in the greater field of systems 
analysis indicates that agent-based modeling enables more flexible simulation 
control, without loss of generality \cite{thatpapermattsent}. Finally, openness 
allows cross-institutional collaboration, increases code robustness, and 
enables the cultivation of an ecosystem of calculation options, 
\cite{softwarecarpentryresource}.  The fundamental concepts of the \Cyclus 
nuclear fuel cycle simulator capture these modern insights so that challenges 
in nuclear fuel cycle analysis can be better addressed.

@@

Historically, the national laboratories have driven development and regulated 
the use of their own tools such as \gls{VISION}\cite{jacobson_verifiable_2010}, 
\gls{DANESS}\cite{van_den_durpel_daness_2009}, 
\gls{DYMOND}\cite{modeling_yacout_2005}, and 
\gls{NFCSim}\cite{schneider_nfcsim:_2005}. 
Outside of the national laboratories, researchers create new nuclear fuel cycle 
simulation tools when existing tools are not available or not sufficiently 
general to calculate their metrics of interest with their method of choice.  
With limited access to the national laboratory tools and a need to customize 
them for research purposes, universities and private industry researchers 
``reinvent the wheel'' by developing tools of their own from scratch and 
tailored to their own needs (e.g., \gls{CAFCA}\cite{guerin_benchmark_2006} and 
\gls{COSI}\cite{cosi} respectively).

@@

System dynamics is a popular approach for modeling nuclear fuel cycles in current
tools \cite{VISION,CAFCA}. However, it has been shown in systems analysis literature
\cite{macal_agent-based_2010} that, formally, the
set of system dynamics models is a strict subset of agent-based models.
That is, any system dynamics model can be translated
into an agent-based model. Furthermore, agent-based
techniques can provide a richer model than can system dynamics. To enable the 
broadest range of simulations in a generic fashion, \gls{ABM} is therefore 
superior to system dynamics.

>>>>

An alternative to these approaches is a dynamic, agent-based simulator with an 
open platform and an ecosystem of research-driven capabilities.  Modern 
insight indicates that dynamic nuclear fuel cycle analysis more realistically 
supports a broader range of simulation goals than static analysis 
\cite{piet_dynamic_2011}. Similarly, experience in the greater field of systems 
analysis indicates that agent-based modeling enables more flexible simulation 
control, without loss of generality \cite{macal_agent-based_2010}. Finally, openness 
allows cross-institutional collaboration, increases code robustness, and 
enables the cultivation of an ecosystem of calculation options, 
\cite{softwarecarpentryresource}.  The fundamental concepts of the \Cyclus 
nuclear fuel cycle simulator capture these modern insights so that challenges 
in nuclear fuel cycle analysis can be better addressed.

@@

Historically, the national laboratories have driven development and regulated 
the use of their own tools such as \gls{VISION}\cite{jacobson_verifiable_2010}, 
\gls{DANESS}\cite{van_den_durpel_daness_2009}, 
\gls{DYMOND}\cite{yacout_modeling_2005}, and 
\gls{NFCSim}\cite{schneider_nfcsim:_2005}. 
Outside of the national laboratories, researchers create new nuclear fuel cycle 
simulation tools when existing tools are not available or not sufficiently 
general to calculate their metrics of interest with their method of choice.  
With limited access to the national laboratory tools and a need to customize 
them for research purposes, universities and private industry researchers 
``reinvent the wheel'' by developing tools of their own from scratch and 
tailored to their own needs (e.g., \gls{CAFCA}\cite{guerin_impact_2009} and 
\gls{COSI}\cite{boucher_cosi_2005,boucher_cosi:_2006,meyer_new_2009,coquelet-pascal_comparison_2011} respectively).

@@

System dynamics is a popular approach for modeling nuclear fuel cycles in current
tools 
\cite{jacobson_vision_2009,van_den_durpel_daness_2009,guerin_impact_2009,guerin_benchmark_2009}. However, it has been shown in systems analysis literature
\cite{macal_agent-based_2010} that, formally, the
set of system dynamics models is a strict subset of agent-based models.
That is, any system dynamics model can be translated
into an agent-based model. Furthermore, agent-based
techniques can provide a richer model than can system dynamics. To enable the 
broadest range of simulations in a generic fashion, \gls{ABM} is therefore 
superior to system dynamics.

### 2014-10-22 fundamentals/qa.tex
reduces the chaos in the bibliographies to just one file of chaos

Simulation science - like experimental science - has since the begining had the 
problem of identifying trustworthy results from the nonsensical and the noise.
This is epitomized in the Charles Babbage quote, ``On two occasions I have been asked, 
`Pray, Mr. Babbage, if you put into the machine wrong figures, will the right 
answers come out?' ... I am not able rightly to apprehend the kind of confusion 
of ideas that could provoke such a question.'' \cite{babbage2011passages}. 
The \emph{garbage in, garbage out} phenomenon is not the only scenario which a
simulator must gaurd against; ensuring correctness is equally important.

Multiple strategies collectively known a \emph{quality assurance} (QA) have 
been invented over the years to mitigate the structural and algorithmic errors
on the part of simulators themselves. These include \emph{verification and validation}
(V\&V) \cite{boehm1989software}, \emph{uncertainty quantification} (UQ) 
\cite{sacks1989design}, testing, and others. V\&V and UQ come from 
a distinctly computational science background while testing \emph{et al.} come from 
a software development bent.

Nuclear enginerering code quality is often governed by NQA-1, an ASME specification 
whose latest revision appeared in 2009 \cite{NQA-1a-2009}. This is primarily 
used for designing reactors. However, it is general enough to apply to 
cyclus. Cyclus has adopted an \emph{agile} development process \cite{larman2004agile}, 
interpreting NQA-1 in a manner similar to NEAMS \cite{neams-qa} or PyNE \cite{pyne-vv}.

@@

Cyclus uses the Google Test framework \cite{gtest} as a harness for running unit 
tests. Sufficient unit tests are required for any proposed change to the cyclus
code base. Currently cyclus implements over 450 unit tests and cycamore has 
85.  These cover approximately 65\% of their respective code bases. This number 
is expected to grow over time.

@@

Intergration tests are any tests that combine multiple elemnents of the 
cyclus interface and test that they work correctly with each other.  By analogy, 
simply because the gears (units) are made correctly does not imply that the 
clock (integration) will run smoothly, run at all, or give the correct time.
In cyclus and cycamore, integration tests are performed by running sample
simulations and verifying that results are what would be expected ahead of 
time. Precanned input files have thus been contsructed. Their results 
are inspected and compared via nose \cite{nosetests}, a Python test framework.
Python will run cyclus as a subprocess for each integration test. In this
way cyclus code units are tested in the full context that they will
executed. This category of testing is especially useful for ensuring that 
major cyclus components are functioning as expected.

@@

All of the public interface (the API) must be documented as part of cyclus
QA policy. This certifies that the intention for a code unit is communicated 
in way that is separate from its implementation. This policy is important 
because implementations are often wrong, hence bugs. They are thus unreliable 
resources about themselves. Truly trustworthy communication about what 
software should be doing comes from the author explicitly stating the 
goals of an API in prose form. Furure development of the code is then able to 
compare the implemetation versus the documentation to discern whether or not they 
are in agreement. Documentation is thus a secondary higher-level information 
stream.  In cyclus, this information is aggregated together into static 
websites with the doxygen \cite{van2008doxygen} and sphinx \cite{brandl2014sphinx} 
tools. These websites are placed online at \url{http://fuelcycle.org}.

@@

Version control is the history-preserving mechanism by which the provenance of 
a code is recored. Cyclus uses a well-established tool called git \cite{git_2014}
for this purpose. Git is currently the most popular utilty in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every 
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
Though there and many praises and criticisms of git, one 
feature that the tool performs exceptionally well is the notion of code branches.
These are distinct pathways in the history that branch out of a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simeltaneous
branches may exist at all point in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an 
accompanying message. Thus 
it is possible for cyclus to accurarely replay the entirety of who did what to the
code when.

Cyclus uses a strategy known as \emph{git flow} \cite{kalliamvakou2014code} 
to manage topical branches where individuals develop code, the mainline development
branch (simply known as develop), and the stable branch (called master).
In addition to these rules and delineation, cyclus also imposes rules on 
when it is accecptable from a quality control standpoint to merge code from 
the personal topical branches into develop. All software at this stage 
must have sufficient and passing tests and reasonable documentation. Without
these core pieces proposed changes are not allowed into cyclus.

The act of proposing a change is known as a \emph{pull request}. The main cyclus is 
hosted remotely on the GitHub website \cite{dabbish2012social}. The online
mechanism for pull requests allows for code review by a member of the cyclus 
core team other than the author prior to any change being included. Non-members
of the cyclus development team are allowed and encouraged to submit and review 
pull requests. However, only members of the cyclus core team are allowed to 
merge in any changes and only after the QA standards have been met.

@@

In any multi-person software project, there is a tension between how individuals
wish to write code. Every person tends to have their own custom style. To remedy this
coding style guides are the software analogy to the natural language ones, 
such as the Chicago Manual of Style. Cyclus strictly enforces the use of the 
Google C++ Style Guide \cite{weinberger2008google} for all software contributions.
This means that all developers of cyclus nominally write cyclus code in the same 
way.  This homogenization may be a hurdle to new developers but ultimately 
makes the it much more legible. Adoption of a single style guide also settles
timeless arguments over what style is `the best.' Having only one means that 
there is a common enemy for all developers to dislike equally.

@@

Continuous integration (CI) the idea that software should be tested and validated 
as it is being developed, rather than as a final stage in a longer development 
cycle.  Under CI every pull request is tested independently immeadiately after 
it is proposed. Such tests shoudl be run on all officially supported platforms. 
Cyclus uses a CI platform called Polyphemus \cite{polyphemus_2014}.

Polyphemus serves as an intermediary between GitHub pull requests on the frontend 
and temporary cyclus servers on the backend. These servers are hosted by 
the Build \& Test Laboratory (BaTLab) \cite{batlab_2014} at the University of 
Wisconsin-Madison. Whenever a pull request is created, Polyphemus performs 
the following actions:

>>>>

Simulation science - like experimental science - has since the begining had the 
problem of identifying trustworthy results from the nonsensical and the noise.
This is epitomized in the Charles Babbage quote, ``On two occasions I have been asked, 
`Pray, Mr. Babbage, if you put into the machine wrong figures, will the right 
answers come out?' ... I am not able rightly to apprehend the kind of confusion 
of ideas that could provoke such a question.'' \cite{babbage_passages_2011}. 
The \emph{garbage in, garbage out} phenomenon is not the only scenario which a
simulator must gaurd against; ensuring correctness is equally important.

Multiple strategies collectively known a \emph{quality assurance} (QA) have 
been invented over the years to mitigate the structural and algorithmic errors
on the part of simulators themselves. These include \emph{verification and validation}
(V\&V) \cite{boehm_software_1989}, \emph{uncertainty quantification} (UQ) 
\cite{sacks_design_1989}, testing, and others. V\&V and UQ come from 
a distinctly computational science background while testing \emph{et al.} come from 
a software development bent.

Nuclear enginerering code quality is often governed by NQA-1, an ASME specification 
whose latest revision appeared in 2009 \cite{asme_nqa-1a-2009_2009}. This is primarily 
used for designing reactors. However, it is general enough to apply to 
cyclus. Cyclus has adopted an \emph{agile} development process 
\cite{larman_agile_2004}, 
interpreting NQA-1 in a manner similar to NEAMS \cite{neams_nuclear_2013} or 
PyNE \cite{biondo_quality_2014}.

@@

Cyclus uses the Google Test framework \cite{inc_googletest_2008} as a harness for running unit 
tests. Sufficient unit tests are required for any proposed change to the cyclus
code base. Currently cyclus implements over 450 unit tests and cycamore has 
85.  These cover approximately 65\% of their respective code bases. This number 
is expected to grow over time.

@@

Intergration tests are any tests that combine multiple elemnents of the 
cyclus interface and test that they work correctly with each other.  By analogy, 
simply because the gears (units) are made correctly does not imply that the 
clock (integration) will run smoothly, run at all, or give the correct time.
In cyclus and cycamore, integration tests are performed by running sample
simulations and verifying that results are what would be expected ahead of 
time. Precanned input files have thus been contsructed. Their results 
are inspected and compared via nose \cite{pellerin_nose_2007}, a Python test framework.
Python will run cyclus as a subprocess for each integration test. In this
way cyclus code units are tested in the full context that they will
executed. This category of testing is especially useful for ensuring that 
major cyclus components are functioning as expected.

@@

All of the public interface (the API) must be documented as part of cyclus
QA policy. This certifies that the intention for a code unit is communicated 
in way that is separate from its implementation. This policy is important 
because implementations are often wrong, hence bugs. They are thus unreliable 
resources about themselves. Truly trustworthy communication about what 
software should be doing comes from the author explicitly stating the 
goals of an API in prose form. Furure development of the code is then able to 
compare the implemetation versus the documentation to discern whether or not they 
are in agreement. Documentation is thus a secondary higher-level information 
stream.  In cyclus, this information is aggregated together into static 
websites with the doxygen \cite{van_heesch_doxygen:_2008} and sphinx 
\cite{brandl_sphinx_2014} 
tools. These websites are placed online at \url{http://fuelcycle.org}.

@@

Version control is the history-preserving mechanism by which the provenance of 
a code is recored. Cyclus uses a well-established tool called git 
\cite{software_freedom_conservancy_git_2014}
for this purpose. Git is currently the most popular utilty in the recent wave
of \emph{distributed version control}. Briefly stated, this means that every 
user has a complete local copy of the \emph{repository} (the version control
term for the collection of all possible histories).
Though there and many praises and criticisms of git, one 
feature that the tool performs exceptionally well is the notion of code branches.
These are distinct pathways in the history that branch out of a mainline source
tree and then may be \emph{merged} back into the mainline. Multiple simeltaneous
branches may exist at all point in time. Every change to the code is recorded
in the history, along with metadata such as the author, a timestamp, and an 
accompanying message. Thus 
it is possible for cyclus to accurarely replay the entirety of who did what to the
code when.

Cyclus uses a strategy known as \emph{git flow} 
\cite{kalliamvakou_code-centric_2014} 
to manage topical branches where individuals develop code, the mainline development
branch (simply known as develop), and the stable branch (called master).
In addition to these rules and delineation, cyclus also imposes rules on 
when it is accecptable from a quality control standpoint to merge code from 
the personal topical branches into develop. All software at this stage 
must have sufficient and passing tests and reasonable documentation. Without
these core pieces proposed changes are not allowed into cyclus.

The act of proposing a change is known as a \emph{pull request}. The main cyclus is 
hosted remotely on the GitHub website \cite{dabbish_social_2012}. The online
mechanism for pull requests allows for code review by a member of the cyclus 
core team other than the author prior to any change being included. Non-members
of the cyclus development team are allowed and encouraged to submit and review 
pull requests. However, only members of the cyclus core team are allowed to 
merge in any changes and only after the QA standards have been met.

@@

In any multi-person software project, there is a tension between how individuals
wish to write code. Every person tends to have their own custom style. To remedy this
coding style guides are the software analogy to the natural language ones, 
such as the Chicago Manual of Style. Cyclus strictly enforces the use of the 
Google C++ Style Guide \cite{weinberger_google_2008} for all software contributions.
This means that all developers of cyclus nominally write cyclus code in the same 
way.  This homogenization may be a hurdle to new developers but ultimately 
makes the it much more legible. Adoption of a single style guide also settles
timeless arguments over what style is `the best.' Having only one means that 
there is a common enemy for all developers to dislike equally.

@@

Continuous integration (CI) the idea that software should be tested and validated 
as it is being developed, rather than as a final stage in a longer development 
cycle.  Under CI every pull request is tested independently immeadiately after 
it is proposed. Such tests shoudl be run on all officially supported platforms. 
Cyclus uses a CI platform called Polyphemus \cite{scopatz_polyphemus_2014}.

Polyphemus serves as an intermediary between GitHub pull requests on the frontend 
and temporary cyclus servers on the backend. These servers are hosted by 
the Build \& Test Laboratory (BaTLab) \cite{uw_batlab_team_batlab_2014} at the University of 
Wisconsin-Madison. Whenever a pull request is created, Polyphemus performs 
the following actions:
